---
title: "Extracting lists from a data frame"
author: "G T Maskall"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---

# Introduction

Whether you've read some data from JSON, or otherwise, you've ended up with these weird
lists in columns. What do you do?

We'll use a snippet of data from the Yelp challenge dataset. We've got some information
on a business level (business_id). One of those columns is a list of attributes for
the business. Now, you're a good data scientist so you'll be itching to extract these
out into columns so you can use them as features for something cool and interesting.

```{r, initalise_data}
attributes <- readRDS("attributes.rds")
```

```{r, load_packages}
library(tidyverse)
```

# Data format

The data look like this:

```{r, data_glimpse}
attributes %>% glimpse
```

In order to focus on the key columns of interest, we'll just grab the business_id
and attributes columns to reduce the clutter. They're the bits we're interested
in and we can always join the data back up later anyway.

```{r, select_cols}
# and key things we're interested in is attributes for business_id:
attributes %>% 
    select(business_id, attributes) 
# let's start testing ideas on the first element
tmp <- attributes$attributes[1]

# extracting successive elements
tmp %>% map_chr(1)
tmp %>% map_chr(2)
tmp %>% map_chr(3)
tmp %>% map_chr(4)
```

Okay, so we might be onto something here. We seem to have been able to extract 
key:value pairs from the list. It's a start! But we don't want to have to
manually specify list index for each row. We want to end up automatically
generating a set of column features from whatever we've got in our list
attributes.

# Getting there

We might first think of unnest. Great, this will do what we want, right?

```{r, unnest_1}
try(
attributes %>% 
    select(business_id, attributes) %>% 
    unnest )
```

Okay, it didn't like that. What went wrong? We're not taking into account the presence
of some empty attribute lists. Can we filter on the length of each? Let's try creating
a length feature

```{r, calc_length_1}
attributes %>% 
    select(business_id, attributes) %>% 
    mutate(Num_attributes = length(attributes))
```

Well they're not all length 10. We know that. We've just calculated the length of
the whole column (10 rows). Let's try again, this time using map_int to return the
expected integers

```{r, calc_length_2}
attributes %>% 
    select(business_id, attributes) %>% 
    mutate(num_atts = map_int(attributes, length)) 
```

Now that's more like it! Let's try unnest again now we can filter out that pesky
null attribute.

```{r, unnest_2}
attributes %>% 
    select(business_id, attributes) %>% 
    mutate(num_atts = map_int(attributes, length)) %>% 
    filter(num_atts > 0) %>% 
    unnest %>% 
    separate(attributes, into = c("key", "value"), extra = "merge") 
```

This is looking much more hopeful! We have successfully extracted the first
attribute key. The final column, the value part, happily doesn't show because
some are fairly long.

All we need to do now is to use the good old spread function from tidyr to give
us our feature columns for our final result!

# Extracting features

```{r, generate_features}
attributes_unnested <- attributes %>% 
    select(business_id, attributes) %>% 
    mutate(num_atts = map_int(attributes, length)) %>% 
    filter(num_atts > 0) %>% 
    unnest %>% 
    separate(attributes, into = c("key", "value"), extra = "merge") %>% 
    spread(key, value) 

attributes_unnested %>% glimpse
```

We can now start to get on with our usual data cleaning. For example, what values do
we have for "WiFi"?

```{r, wifi}
attributes_unnested %>%
    count(WiFi)
```

# Further work

As you can see, there are still some features we might want to extract further, such as
Ambience. This is for another day, but you might decide where you want to go with that
is to split the values out into yet more columns having, perhaps, prepended "Ambience:"
to each subkey to generate "Ambience:romantic" and "Ambience:intimate" features, for example.
